\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

---
title: "Factors influencing Turkiye's Growth"
author: "Fei Wu, Merve Onal, Siyao Chen, Ye Khaung Oo"
date: "2024-04-23"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE, tidy=TRUE}

#Ye Working on this

#Checking working directory
getwd()

#Loading packages
library(tidyverse); library(lubridate);library(zoo);
library(trend); library(ggplot2); library(ggthemes);
library(readxl); library(here); library(dplyr);

#importing dataset

#setting theme
theme_set(theme_minimal())

```

```{r, message = FALSE}

# Here is the list of file names to import
#file_names <- c(
  "Data_for_R_Project.xlsx"
#)

```
```{r}

# Set the directory path
#directory_path <- "~/Chen_Onal_Oo_Wu/Data"

# Define the file name
#file_name <- "Data_for_R_Project.xlsx"

# Construct the full file path
#file_path <- file.path(directory_path, file_name)

# Read the Excel file into a data frame
#if (file.exists(file_path)) {
  #data <- read_excel(file_path)
  # Convert "Years" column to Date class
  #data <- mutate(data, Years = as.Date(Years, format = "%Y"))
#} else {
 # message("File", file_name, "not found.")
#}

```

```{r}

# Read the Excel file into a data frame
#if (file.exists(file_path)) {
  #data <- read_excel(file_path)
  # Convert "Years" column to Date class
  #data <- mutate(data, Years = as.Date(Years, format = "%Y"))
#} else {
  #message("File", file_name, "not found.")
#}

```

```{r}
# Here I create a sequence of dates from 1999 to 2022
# Define the start and end years
#start_year <- 1999
#end_year <- 2022

# Create a sequence of years from start_year to end_year
#year_sequence <- seq(start_year, end_year, by = 1)

# Print the sequence of years
#year_sequence


```

# Rationale and Research Questions

Does electrcity consumption increase the GDP growth 

## Background

We analyze the effect of investment growth which corresponds to capital accumulation and population growth on the GDP growth of Turkiye for the 1999-2022 period. We also analyze the impact of electrcity consumption on the GDP growth.

## Significance

Understanding the relationship between electrcity consumption and GDP growth is crucial. Understanding the behavior of electrcity consumption in relation to the economy is very important for improve
a stable economic growth and development (Kasperowicz, 2014: 46).


## Theoretical Context

The hypothesis underlying this research is that an increase in electrcity consumption leads to GDP growth. This hypothesis is grounded in the understanding that investment, population and electrcity consumption which explains GDP growth.

## Visualize
```{r}

#Merve Working on This
setwd("~/Chen_Onal_Oo_Wu/Data/Processed")
data <- read_excel("Data_for_R_Project.xlsx")
str(data)
data$Years <- as.Date(data$Years)
install.packages("Rmisc")
library(Rmisc)
# Function to create time series plot with custom title and axis labels
create_ts_plot <- function(data, variable_name, main_title, xlab_title, ylab_title) {
  plot(data, main = main_title, xlab = xlab_title, ylab = ylab_title)
}
  # Remove rows with missing values in any of the columns
  data <- na.omit(data)
  
library(ggplot2)
# Assuming you have a data frame named 'my_data' containing columns 'Invgro', 'GDPgro', and 'Popgro'
# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Invgro)) +
  geom_line(color = "blue") 
# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Popgro)) +
  geom_line(color = "red") 
 
# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Electirictygro)) +
  geom_line(color = "green") 
# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Electirictygro)) +
  geom_line(color = "black") 
# Create the plot using ggplot2
ggplot(data, aes(x = Popgro, y = Electirictygro)) +
  geom_line(color = "yellow") 
# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Popgro)) +
  geom_line(color = "purple")
ggplot(data, aes(Years)) + 
  geom_line(aes(y = Invgro, colour = "red")) + 
  geom_line(aes(y = Popgro, colour = "blue")) +
  geom_line(aes(y = Electirictygro, colour = "green"))+
          ylab('Values')+xlab('date')
 

```

>Explanation: From the plot, we notice a slightly decreasing trend throughout the 24 years period from 1999 to 2022. Furthermore, we also can notice 

# Dataset Information



\newpage

# Exploratory Analysis 
```{r}
#Merve Working on t his
setwd("~/Chen_Onal_Oo_Wu/Data/Processed")
data <- read_excel("Data_for_R_Project.xlsx")
str(data)
data$Years <- as.Date(data$Years)
ts_data <- ts(data$GDPgro, start = min(1999), end = max(2022), frequency = 1)
plot(ts_data, main = "GDP Growth Over Time", xlab = "Years", ylab = "GDP Growth")
plot(ts_data, main = "Investment Growth Over Time", xlab = "Years", ylab = "Investment growth")
plot(ts_data, main = "Population Growth Over Time", xlab = "Years", ylab = "Population growth")
plot(ts_data, main = "Electrcity Consumption Growth Over Time", xlab = "Years", ylab = "Electrcity consumption growth")

install.packages("Rmisc")
library(Rmisc)
# Function to create time series plot with custom title and axis labels
create_ts_plot <- function(data, variable_name, main_title, xlab_title, ylab_title) {
  plot(data, main = main_title, xlab = xlab_title, ylab = ylab_title)
}


  # Remove rows with missing values in any of the columns
  data <- na.omit(data)
  
library(ggplot2)

# Assuming you have a data frame named 'my_data' containing columns 'Invgro', 'GDPgro', and 'Popgro'

# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Invgro)) +
  geom_line(color = "blue") 

# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Popgro)) +
  geom_line(color = "red") 
 
# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Electirictygro)) +
  geom_line(color = "green") 

# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Electirictygro)) +
  geom_line(color = "black") 

# Create the plot using ggplot2
ggplot(data, aes(x = Popgro, y = Electirictygro)) +
  geom_line(color = "yellow") 

# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Popgro)) +
  geom_line(color = "purple")

ggplot(data, aes(Years)) + 
  geom_line(aes(y = Invgro, colour = "red")) + 
  geom_line(aes(y = Popgro, colour = "blue")) +
  geom_line(aes(y = Electirictygro, colour = "green"))+
          ylab('Values')+xlab('date')
 


\newpage
## Visualize
```{r}
setwd("~/Chen_Onal_Oo_Wu/Data/Processed")
data <- read_excel("Data_for_R_Project.xlsx")
str(data)
data$Years <- as.Date(data$Years)

install.packages("Rmisc")
library(Rmisc)
# Function to create time series plot with custom title and axis labels
create_ts_plot <- function(data, variable_name, main_title, xlab_title, ylab_title) {
  plot(data, main = main_title, xlab = xlab_title, ylab = ylab_title)
}


  # Remove rows with missing values in any of the columns
  data <- na.omit(data)
  
library(ggplot2)

# Assuming you have a data frame named 'my_data' containing columns 'Invgro', 'GDPgro', and 'Popgro'

# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Invgro)) +
  geom_line(color = "blue") 

# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Popgro)) +
  geom_line(color = "red") 
 
# Create the plot using ggplot2
ggplot(data, aes(x = GDPgro, y = Electirictygro)) +
  geom_line(color = "green") 

# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Electirictygro)) +
  geom_line(color = "black") 

# Create the plot using ggplot2
ggplot(data, aes(x = Popgro, y = Electirictygro)) +
  geom_line(color = "yellow") 

# Create the plot using ggplot2
ggplot(data, aes(x = Invgro, y = Popgro)) +
  geom_line(color = "purple")

ggplot(data, aes(Years)) + 
  geom_line(aes(y = Invgro, colour = "red")) + 
  geom_line(aes(y = Popgro, colour = "blue")) +
  geom_line(aes(y = Electirictygro, colour = "green"))+
          ylab('Values')+xlab('date')
 

```
## Time Series Analysis
```{r}
#Merve Working On This

# creating time series for GDP data
ts_GDPdata <- ts(data$GDPgro, start = min(1999), end = max(2022), frequency = 24)

# creating time series for Investment data
ts_Invdata <- ts(data$Invgro, start = min(1999), end = max(2022), frequency = 24)

# creating time series for Populaton data
ts_Popdata <- ts(data$Popgro, start = min(1999), end = max(2022), frequency = 24)

# creating time series for Electricity data
ts_Elecdata <- ts(data$Electirictygro, start = min(1999), end = max(2022), frequency = 24)

# Decomposing the time series
ts_GDPdata.decomp <- stl(ts_GDPdata, s.window = "periodic")
plot(ts_GDPdata.decomp)
ts_Invdata.decomp <- stl(ts_Invdata, s.window = "periodic")
plot(ts_Invdata.decomp)
ts_Popdata.decomp <- stl(ts_Popdata, s.window = "periodic")
plot(ts_Popdata.decomp)
ts_Elecdata.decomp <- stl(ts_Elecdata, s.window = "periodic")
plot(ts_Elecdata.decomp)



```
# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>
Has electrcity consumption contributed to GDP growth during the period from 1999 to 2022 in Turkiye?

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
