\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

---
title: "Time Series"
author: "Merve Onal"
date: "2024-04-23"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}

#Check your working directory
#Load the tidyverse, lubridate, zoo, and trend packages
#Set your ggplot theme

getwd()
#install.packages
library(tidyverse)
library(lubridate)
library(zoo)
library(trend)
library(readxl)
library(here)
library(foreign)
library(dplyr)

# Here I set the ggplot theme
theme_set(theme_minimal())

```

```{r, message = FALSE}

#Set your ggplot theme

setwd('~/Chen_Onal_Oo_Wu/Data')

# Here is the list of file names to import
file_names <- c(
  "Data_for_R_Project.xlsx"
)

```
```{r}

# Set the directory path
directory_path <- "~/Chen_Onal_Oo_Wu/Data"

# Define the file name
file_name <- "Data_for_R_Project.xlsx"

# Construct the full file path
file_path <- file.path(directory_path, file_name)

# Read the Excel file into a data frame
if (file.exists(file_path)) {
  data <- read_excel(file_path)
  # Convert "Years" column to Date class
  data <- mutate(data, Years = as.Date(Years, format = "%Y"))
} else {
  message("File", file_name, "not found.")
}

```

```{r}

# Read the Excel file into a data frame
if (file.exists(file_path)) {
  data <- read_excel(file_path)
  # Convert "Years" column to Date class
  data <- mutate(data, Years = as.Date(Years, format = "%Y"))
} else {
  message("File", file_name, "not found.")
}

```

```{r}
# Here I create a sequence of dates from 1999 to 2022
# Define the start and end years
start_year <- 1999
end_year <- 2022

# Create a sequence of years from start_year to end_year
year_sequence <- seq(start_year, end_year, by = 1)

# Print the sequence of years
year_sequence


```

# Rationale and Research Questions

## Background

here relationship between electricity consumption and gdp growth

## Significance

Understanding the relationship between electricity consumption and GDP growth is crucial. Understanding the behavior of electricity consumption in relation to the economy is very important for improve
a stable economic growth and development (Kasperowicz, 2014: 46).


## Theoretical Context

The hypothesis underlying this research is that an increase in electricity consumption leads to GDP growth. This hypothesis is grounded in the understanding that investment, population and electricity consumption whixh explains GDP growth.

## Visualize
```{r}
library(ggplot2)


# Create the line plot

gdp_plot <- ggplot(data = Data_for_R_Project, aes(x = Years)) +
  geom_line(aes(y = GDPgro), color = "blue", linetype = "solid", linewidth = 1) +
  geom_line(aes(y = Popgro), color = "red", linetype = "dashed", linewidth = 1) +
  geom_line(aes(y = Invgro), color = "green", linetype = "dotted", linewidth = 1) +
  geom_line(aes(y = Electirictygro), color = "orange", linetype = "dotdash", linewidth = 1) +
  labs(title = "Trends Over Time",
       x = "Years",
       y = "Values") +
   scale_y_continuous(labels = scales::comma,  
                     limits = c(0, 10)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# Print the plot
# Print the plot
print(gdp_plot) +
  geom_line(aes(y = GDPgro), color = "blue", linetype = "solid", size = 1) +
  geom_line(aes(y = Popgro), color = "red", linetype = "dashed", size = 1) +
  geom_line(aes(y = Invgro), color = "green", linetype = "dotted", size = 1) +
  geom_line(aes(y = Electirictygro), color = "orange", linetype = "dotdash", size = 1) +
  labs(title = "Trends Over Time",
       x = "Years",
       y = "Values") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





```

>Explanation: From the plot, we notice a slightly decreasing trend throughout the 24 years period from 1999 to 2022. Furthermore, we also can notice 

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
